---
title: "Homework 3"
output: github_document
---

```{r setup, include= FALSE}
library(tidyverse)
library(hexbin)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1 

#### How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


#### Creating a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


#### Creating a table

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


#### Apples vs ice cream

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


### Problem 2 

#### Load, tidy, and describe the accelerometer data set. 
```{r}
accel_df = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
    ) %>%
  mutate(
    minute = as.numeric(minute),
    day = as.factor(day),
    day = fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
    weekend = as.numeric(day %in% c("Saturday", "Sunday")),
    day_type = factor(recode(weekend, '1' = "weekend", '0' = "weekday"))) %>%
  select(-weekend)
    
  
    
  
```
#### data description


#### Creating total activity per day variable and associated table.
```{r}
test =
accel_df %>%
  group_by(day, week) %>%
  summarize(total_activity = sum(activity_count)) %>%
  pivot_wider(
    names_from = day,
    values_from = total_activity)
 
test %>%
  knitr::kable()

```


#### Creating a single-panel plot that shows the 24-hour activity time courses for each day.
```{r}
activity_plot =
accel_df %>%
  ggplot(aes(x = minute, y = activity_count, color = day)) + 
  geom_line(aes(group = day))
  geom_smooth(aes(group = day), se = FALSE)
  labs(
    title = "24 hr Activity Time Courses"
    
  )
 


```


#### Description of plot



### Problem 3

```{r setup_2, include = FALSE}
data("ny_noaa")
```

##### Data exploration and description

##### Data cleaning
```{r}
ny_noaa_clean = 
 ny_noaa %>%
 separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
 mutate(
   prcp = prcp/10,
   year = as.numeric(year),
   month = as.numeric(month),
   day = as.numeric(day),
   tmax = as.numeric(tmax)/10,
   tmin = as.numeric(tmin)/10
 )
```

#### Most commmonly observed snowfall value
```{r}
common_snow = 
ny_noaa_clean %>%
  group_by(snow) %>%
  summarize(n_obs = n()) 


```


#### Max temp in January and July
```{r}
 
ny_noaa_clean %>%
filter(month %in% c("1", "7")) %>%
mutate(month = month.name[month]) %>%
group_by(id, year, month) %>%
summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>%
ggplot(aes(x = year, y = mean_tmax, group = id, color = id)) +
  geom_line() +
  geom_path() +
  facet_grid(~ month) +
  labs(title = "Average Temp for January and July", x = "Year", y = "Average Maximum Temp (C)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))


```

#### Two-panel plot
```{r}
tmaxmin =
  ny_noaa_clean %>%
  drop_na(tmax, tmin) %>%
  ggplot(aes(x = tmin, y = tmax)) +
  labs(
    title = "Temperature (max vs. min) Hexplot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)") +
  geom_hex() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.text = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  theme(legend.box = "horizontal")
  

snow_dist = 
  ny_noaa_clean %>%
  drop_na(snow) %>%
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = year, y = snow)) +
  labs(
    title = "Snowfall Values Boxplot",
    x = "Year",
    y = "Snowfall (mm)") +
  geom_boxplot(aes(group = year)) +
  theme(plot.title = element_text(hjust = 0.5))
  
     
tmaxmin + snow_dist
  

```

